Testing the Evolution of the Correlations between Supermassive Black Holes and their Host Galaxies using Eight Strongly Lensed Quasars

### 2020-09-20 manuscript/manuscript.tex
Final update to the paper, after the final review.

{\bf WFI2033} ~ 
{\color{blue}
%In addition to the F160W filter, images through the four filters F125W, F140W, F555W, F814W are also available in the \hst\ archive. As for HE0435, we infer the host color in the image plane. We find a stellar population with 0.625 Gyr is a good match to the colors.
In addition to WFC3/F160W, WFI2033 has been observed by \hst\ in  four bands, i.e. WFC3/F125W (GO-12874; PI: D. Floyd), WFC3/F140W (GO-13732; PI: A. Nierenberg), ACS/F555W+F814W  (GO-9744; PI: C. S. Kochanek). Similar to HE0435, the lens modelling and the host photometry of WFI2033 have been inferred through these four bands. The results are presented in Figure~\ref{fig:app_WFI2033}-(a)$-$(d). Unfortunately, the  F125W data are too shallow to robustly detect the host.  Thus, we do not use the F125W band for the SED fitting. Using four-band photometry, we find a stellar population of 0.625 Gyr age provides a good match to the colors, as shown in Figure~\ref{fig:app_WFI2033}-(e).
\begin{figure*}
\centering
\subfloat[WFI2033 F125W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f125w_inference.pdf}}\\
\subfloat[WFI2033 F140W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f140w_inference.pdf}}\\
\subfloat[WFI2033 F555W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f555w_inference.pdf}}\\
\subfloat[WFI2033 F814W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f814w_inference.pdf}}\\
\subfloat[Host galaxy stellar template inference]{\includegraphics[ width=0.5\textwidth]{fig/WFI2033_host_color.pdf}}
\caption{\label{fig:app_WFI2033} 
Same as Figure~\ref{fig:app_WFI2033} but for WFI2033. The F125W dare are too shallow to detect the host and are thus omitted in the fit.}
\end{figure*} 
}

{\bf HE1104 and the other systems} ~ HE1104 has also been observed by \hst\ through the F555W and F814W bands. However, given the limited exposure time in these two bands, the lensed arcs are too faint to be detected; thus, we are not able to infer the color of the host for HE1104. No multi-band information is available for the other systems. Thus, for these cases, we follow D20 and assume a typical stellar population age, i.e. 1 Gyr and 0.625 Gyr for systems as $z<1.44$ and $z>1.44$, respectively. Of course, this choice is not unique. However, we stress that as we are mostly interested in comparison with D20 this strategy is meant to minimize differences between the two approaches.

@@

Considering the simulations by~\citet{Ding2017a} and the fact that we are able to obtain very consistent host magnitudes for the HE0435, RXJ1131 and WFI2033 using the independent approaches, the fidelity of the inferred magnitude is expected to be well characterized by the quoted uncertainties.
\blue{
Our systems do not have sufficient multi-band information to constrain their metallicity and star formation histories. Therefore, we adopted a simple stellar population template to calculate \mstar, but reflect the lack of information in the associated uncertainty. Following \citet{Bell2001}, we estimate a typical uncertainty of 0.2~dex for all the \mstar\ measurements in this paper.}

\subsection{The \mbh-\mstar\ relation}\label{sec:relation}
In Figure~\ref{fig:scaling_relation}-(left), we plot \mbh\ vs \mstar\ for our sample, together with the comparison sample introduced in Section~\ref{sec:sample_select}.  We expect that the uncertainty of the \mbh\ (i.e., $0.4~$dex) dominates the error budget for the entire sample. Following D20, we adopt, as {\bf a} baseline, a local \mbh-\mstar\ of the form:
\begin{eqnarray}
\label{eq:MMlocal}
\log \big( \frac{\mathcal M_{\rm BH}}{10^{7}M_{\odot}})= \alpha + \beta \log(\frac{M_*}{10^{10}M_{\odot}}),
  \end {eqnarray}
with  $\alpha = 0.27$, $\beta = 0.98$ based on the local sample of 55 objects measured in a consistent manner. We find that our lensed systems are above the local relation, consistent with the inference from the 32  non-lensed AGNs published by D20 in a  similar redshift range. To quantify the evolution as a function of redshift, we parameterize it as:
\begin{eqnarray}
\label{eq:offset}
\Delta \log \mathcal M_{\rm BH}= \gamma \log (1 + z),
\end{eqnarray} 
where $\Delta \log \mathcal M_{\rm BH}$ is the offset of \mbh\ with respect to the local baseline at fixed \mstar. To make a direct comparison, we reproduce the plot shown in Figure~8 of D20. Then, we add our new lensed AGN measurements and show the result in Figure~\ref{fig:scaling_relation}-(right). We find that the offset from the local relationship is similar for the two samples. Based on Equation~\eqref{eq:offset}, we fit the evolution of the eight lensed systems and obtain $\gamma=1.05\pm0.44$, which agrees with the results of D20 ($\gamma=1.03\pm0.25$ using 32 AGN) within $1-\sigma$ level. The consistency between the two measurements provides an important verification of the accuracy of the results and strengthens the conclusions drawn by D20 that the {\it observed} value of \mbh\ at a fixed \mstar\ tends to be larger at higher redshift than the ones in the local universe.

\blue{
Note that this apparent evolution is obtained directly using the observed sample, before considering selection effects~\citep{Tre++07,Schulze2011,Bennert++2011, Schulze2014,Park15}. For instance, adopting the framework introduced in~\citet{Schulze2015}, D20 estimated that accounting for selection biases would yield a more modest evolution $\gamma=0.50\pm0.25$. The lensed AGN systems considered in this work were selected for time-delay cosmography, based on the availability of a time delay and the known detectability of the host galaxy. This is a complex selection function, different than the one used by D20. Although it is encouraging that the two samples present the same apparent evolution, inferring the true underlying evolution requires modelling the selection function, as studied by D20. A full modelling of the complex selection function is not warranted given the small size of the lensed quasars sample. Thus, at this stage, the lensed quasars should be considered as a check on possible systematic measurement error in the D20 analysis rather than a stand-alone measurement.}

@@

\blue{
We consider all random and systematic effects in Appendix~\ref{sec:diss}. To summarize, the uncertainty level of the \mstar\ in this work is assumed as 0.2~dex  as mentioned in the previous section. The uncertainty in the black hole mass estimates is at a level of 0.4~dex, which dominates the overall error budget.
}

@@

We directly compare our sample to the recent measurements by~\citet[][D20]{Ding2020a}, who used the same approach to derive the \mstar\ and calibrated the \mbh\ using consistent recipes. The \mbh-\mstar\ correlation and its evolution with cosmic time {\bf are} in excellent agreement with the results obtained by D20, as shown in Figure~\ref{fig:scaling_relation}. \blue{Currently, the sample of lensed AGN (i.e., eight systems) is quite limited in size; as a result, given the statistical errors it is not worth modelling selection effect.} Nevertheless, the good agreement with the D20 sample strengthens their conclusions, which can be summarized as follows. First, the growth of the supermassive black hole \blue{somewhat} predates that of its host galaxy during their co-evolution, even when considering total stellar mass, as it is often done at high redshift. However, the actual morphologies of these hosts are likely to be more complex, including a disk and bulge component, as in the case of  RXJ1131. In contrast, the galaxies in the local sample are typically bulge-dominated and the bulge mass is adopted as their  \mstar. Thus, the reported evolution is weaker than what would be inferred by comparing \mbh\ to bulge \mstar\ at all redshifts \citep{Bennert++2011}. Taken together, these results are consistent with a scenario in which the stellar mass is transferred from disk to the bulge at a faster rate than the growth of \mbh\ since $z\sim2$.

@@

\section*{Acknowledgements}
{\bf The authors thank the anonymous referee for helpful suggestions and comments which improved this paper.}
We are grateful to Frederic Courbin, Leon Koopmans for useful comments
and suggestions that improved this manuscript. We thank John Silverman and Vardha Bennert for many conversations on the topic of galaxy and black hole co-evolution.

This work is based in part on observations made with the NASA/ESA Hubble Space Telescope, obtained at the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555. X.D., S.B., and T.T. acknowledge support by the Packard Foundation through a Packard Research fellowship to T.T. This work is supported by JSPS KAKENHI Grant Number JP18H01251 and the World Premier International Research Center Initiative (WPI), MEXT, Japan. This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 787886).
{\bf C.D.F. acknowledge support for this work from the National Science Foundation under Grant Numbers AST-1312329 and AST-1907396.}
SHS thanks the Max Planck Society for support through the Max Planck Research Group.

@@

@@

{\color{blue}
\section{Photometry Inference}\label{sec:photometry}
We describe the details of the fitting for each system and present the inference of the photometry of the host galaxy.

\subsection{HE0435}\label{subsec:HE0435}
We follow the approach described in the previous section. We select 5 isolated stars in this field as initial PSFs to input to the fitting, for a total of 60 fits. Based on the top-eight choices, we perform the weighting algorithm and measure the host flux, host-to-total flux ratio, effective radius and \sersic\ index. The inference results are shown in Table~\ref{tab:host_measure}, (2)$-$(6) columns.

The inference by the best-fit lens model for HE0435 is shown in Figure~\ref{fig:image_inference}-(a). Not surprisingly, we note that the residuals level in the normalized plot appears to be larger than the ones presented by~\citet{Wong2017}. This is primarily due to the fact that the surface brightness of the host galaxy in our model is defined as a \sersic\ profile, which is relatively simple and smooth compared to the pixellated reconstruction technique adopted by {\sc Glee}~\citep{Suy++06, Halk2008, S+H10, Suyu2012} or the shapelet technique adopted by \lenstronomy~\citep{Birrer2015}. The smooth features of \sersic\ profile can not capture the clumps in the host galaxy, such as the star-forming regions. Nevertheless, our approach is sufficient to derive a self-consistent one-step inference of the global host light in terms of the \sersic\ flux, i.e. the quantity commonly measured in the literature for non-lensed samples.

As a cross-check, we compare our host magnitude to the inference of HE0435 by~\citet{Ding2017b}. They inferred the \sersic\ magnitude by fitting to the pixelized host galaxy as reconstructed by~\citet{Wong2017}. The host magnitude measured by~\citet{Ding2017b} is: $mag = 21.75 \pm 0.13$. %, $R_{eff} = 0.82 \pm 0.14$ and $n = 3.94 \pm 0.14$. 
The results are in excellent agreement with the measurement reported here, i.e. $mag = 21.50 \pm 0.35$.

@@

%In summary, considering all random and systematic effects, the uncertainty of \mstar\ relative to that measured by D20 is smaller than the estimated uncertainty in the black hole mass (0.4~dex), which dominates the overall error budget.
}

>>>>

{\bf WFI2033} ~ 
%In addition to the F160W filter, images through the four filters F125W, F140W, F555W, F814W are also available in the \hst\ archive. As for HE0435, we infer the host color in the image plane. We find a stellar population with 0.625 Gyr is a good match to the colors.
In addition to WFC3/F160W, WFI2033 has been observed by \hst\ in  four bands, i.e., WFC3/F125W (GO-12874; PI: D. Floyd), WFC3/F140W (GO-13732; PI: A. Nierenberg), ACS/F555W+F814W  (GO-9744; PI: C. S. Kochanek). Similar to HE0435, the lens modelling and the host photometry of WFI2033 have been inferred through these four bands. The results are presented in Figure~\ref{fig:app_WFI2033}-(a)$-$(d). Unfortunately, the  F125W data are too shallow to robustly detect the host.  Thus, we do not use the F125W band for the SED fitting. Using four-band photometry, we find a stellar population of 0.625 Gyr age provides a good match to the colors, as shown in Figure~\ref{fig:app_WFI2033}-(e).
\begin{figure*}
\centering
\subfloat[WFI2033 F125W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f125w_inference.pdf}}\\
\subfloat[WFI2033 F140W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f140w_inference.pdf}}\\
\subfloat[WFI2033 F555W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f555w_inference.pdf}}\\
\subfloat[WFI2033 F814W]{\includegraphics[ width=0.8\textwidth]{fig/WFI2033_f814w_inference.pdf}}\\
\subfloat[Host galaxy stellar template inference]{\includegraphics[ width=0.5\textwidth]{fig/WFI2033_host_color.pdf}}
\caption{\label{fig:app_WFI2033} 
Same as Figure~\ref{fig:app_WFI2033} but for WFI2033. The F125W dare are too shallow to detect the host and are thus omitted in the fit.}
\end{figure*}

{\bf HE1104 and the other systems} ~ HE1104 has also been observed by \hst\ through the F555W and F814W bands. However, given the limited exposure time in these two bands, the lensed arcs are too faint to be detected; thus, we are not able to infer the color of the host for HE1104. No multi-band information is available for the other systems. Thus, for these cases, we follow D20 and assume a typical stellar population age, i.e., 1 Gyr and 0.625 Gyr for systems as $z<1.44$ and $z>1.44$, respectively. Of course, this choice is not unique. However, we stress that as we are mostly interested in comparison with D20 this strategy is meant to minimize differences between the two approaches.

@@

Considering the simulations by~\citet{Ding2017a} and the fact that we are able to obtain very consistent host magnitudes for the HE0435, RXJ1131 and WFI2033 using the independent approaches, the fidelity of the inferred magnitude is expected to be well characterized by the quoted uncertainties.
Our systems do not have sufficient multi-band information to constrain their metallicity and star formation histories. Therefore, we adopt a simple stellar population template to calculate \mstar, but reflect the lack of information in the associated uncertainty. Following \citet{Bell2001}, we estimate a typical uncertainty of 0.2~dex for all the \mstar\ measurements in this paper.

\subsection{The \mbh-\mstar\ relation}\label{sec:relation}
In Figure~\ref{fig:scaling_relation}-(left), we plot \mbh\ vs \mstar\ for our sample, together with the comparison sample introduced in Section~\ref{sec:sample_select}.  We expect that the uncertainty of the \mbh\ (i.e., $0.4~$dex) dominates the error budget for the entire sample. Following D20, we adopt, as a baseline, a local \mbh-\mstar\ of the form:
\begin{eqnarray}
\label{eq:MMlocal}
\log \big( \frac{\mathcal M_{\rm BH}}{10^{7}M_{\odot}})= \alpha + \beta \log(\frac{M_*}{10^{10}M_{\odot}}),
  \end {eqnarray}
with  $\alpha = 0.27$, $\beta = 0.98$ based on the local sample of 55 objects measured in a consistent manner. We find that our lensed systems are above the local relation, consistent with the inference from the 32  non-lensed AGNs published by D20 in a  similar redshift range. To quantify the evolution as a function of redshift, we parameterize it as:
\begin{eqnarray}
\label{eq:offset}
\Delta \log \mathcal M_{\rm BH}= \gamma \log (1 + z),
\end{eqnarray} 
where $\Delta \log \mathcal M_{\rm BH}$ is the offset of \mbh\ with respect to the local baseline at fixed \mstar. To make a direct comparison, we reproduce the plot shown in Figure~8 of D20. Then, we add our new lensed AGN measurements and show the result in Figure~\ref{fig:scaling_relation}-(right). We find that the offset from the local relationship is similar for the two samples. Based on Equation~\eqref{eq:offset}, we fit the evolution of the eight lensed systems and obtain $\gamma=1.05\pm0.44$, which agrees with the results of D20 ($\gamma=1.03\pm0.25$ using 32 AGN) within $1-\sigma$ level. The consistency between the two measurements provides an important verification of the accuracy of the results and strengthens the conclusions drawn by D20 that the {\it observed} value of \mbh\ at a fixed \mstar\ tends to be larger at higher redshift than the ones in the local universe. {\bf It is possible that the uncertainty of the \mstar\ could be higher than our estimation (i.e., 0.2~dex), given that the star formation histories could vary widely at high redshift ($z\sim$2). However, even when our stellar mass uncertainty is boosted, the evolution inferred using the eight lensed systems is still robust --- $\gamma=1.05\pm0.48$ assuming  $\Delta$\mstar = 0.3~dex, and $\gamma=1.03\pm0.53$ assuming $\Delta$\mstar = 0.4~dex.}

Note that this apparent evolution is obtained directly using the observed sample, before considering selection effects~\citep{Tre++07,Schulze2011,Bennert++2011, Schulze2014,Park15}. For instance, adopting the framework introduced in~\citet{Schulze2015}, D20 estimated that accounting for selection biases would yield a more modest evolution $\gamma=0.50\pm0.25$. The lensed AGN systems considered in this work were selected for time-delay cosmography, based on the availability of a time delay and the known detectability of the host galaxy. This is a complex selection function, different than the one used by D20. Although it is encouraging that the two samples present the same apparent evolution, inferring the true underlying evolution requires modelling the selection function, as studied by D20. A full modelling of the complex selection function is not warranted given the small size of the lensed quasars sample. Thus, at this stage, the lensed quasars should be considered as a check on possible systematic measurement error in the D20 analysis rather than a stand-alone measurement.

@@

We consider all random and systematic effects in Appendix~\ref{sec:diss}. To summarize, the uncertainty level of the \mstar\ in this work is assumed as 0.2~dex  as mentioned in the previous section. The uncertainty in the black hole mass estimates is at a level of 0.4~dex, which dominates the overall error budget.

@@

We directly compare our sample to the recent measurements by~\citet[][D20]{Ding2020a}, who used the same approach to derive the \mstar\ and calibrated the \mbh\ using consistent recipes. The \mbh-\mstar\ correlation and its evolution with cosmic time are in excellent agreement with the results obtained by D20, as shown in Figure~\ref{fig:scaling_relation}. Currently, the sample of lensed AGN (i.e., eight systems) is quite limited in size; as a result, given the statistical errors it is not worth modelling selection effect. Nevertheless, the good agreement with the D20 sample strengthens their conclusions, which can be summarized as follows. First, the growth of the supermassive black hole somewhat predates that of its host galaxy during their co-evolution, even when considering total stellar mass, as it is often done at high redshift. However, the actual morphologies of these hosts are likely to be more complex, including a disk and bulge component, as in the case of  RXJ1131. In contrast, the galaxies in the local sample are typically bulge-dominated and the bulge mass is adopted as their  \mstar. Thus, the reported evolution is weaker than what would be inferred by comparing \mbh\ to bulge \mstar\ at all redshifts \citep{Bennert++2011}. Taken together, these results are consistent with a scenario in which the stellar mass is transferred from disk to the bulge at a faster rate than the growth of \mbh\ since $z\sim2$.

@@

\section*{Acknowledgements}
The authors thank the anonymous referee for helpful suggestions and comments which improved this paper.
We are grateful to Frederic Courbin, Leon Koopmans for useful comments
and suggestions that improved this manuscript. We thank John Silverman and Vardha Bennert for many conversations on the topic of galaxy and black hole co-evolution.

This work is based in part on observations made with the NASA/ESA Hubble Space Telescope, obtained at the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., under NASA contract NAS 5-26555. X.D., S.B., and T.T. acknowledge support by the Packard Foundation through a Packard Research fellowship to T.T. This work is supported by JSPS KAKENHI Grant Number JP18H01251 and the World Premier International Research Center Initiative (WPI), MEXT, Japan. This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement No 787886).
C.D.F. acknowledge support for this work from the National Science Foundation under Grant Numbers AST-1312329 and AST-1907396.
SHS thanks the Max Planck Society for support through the Max Planck Research Group.

@@

\section*{Data Availability}
The data underlying this article are available in the article and in its online supplementary material.

@@

\section{Photometry Inference}\label{sec:photometry}
We describe the details of the fitting for each system and present the inference of the photometry of the host galaxy.

\subsection{HE0435}\label{subsec:HE0435}
We follow the approach described in Section~\ref{sec:Modelling}. We select 5 isolated stars in this field as initial PSFs to input to the fitting, for a total of 60 fits. Based on the top-eight choices, we perform the weighting algorithm and measure the host flux, host-to-total flux ratio, effective radius and \sersic\ index. The inference results are shown in Table~\ref{tab:host_measure}, (2)$-$(6) columns.

The inference by the best-fit lens model for HE0435 is shown in Figure~\ref{fig:image_inference}-(a). Not surprisingly, we note that the residuals level in the normalized plot appears to be larger than the ones presented by~\citet{Wong2017}. This is primarily due to the fact that the surface brightness of the host galaxy in our model is defined as a \sersic\ profile, which is relatively simple and smooth compared to the pixellated reconstruction technique adopted by {\sc Glee}~\citep{Suy++06, Halk2008, S+H10, Suyu2012} or the shapelet technique adopted by \lenstronomy~\citep{Birrer2015}. The smooth features of \sersic\ profile can not capture the clumps in the host galaxy, such as the star-forming regions. Nevertheless, our approach is sufficient to derive a self-consistent one-step inference of the global host light in terms of the \sersic\ flux, i.e., the quantity commonly measured in the literature for non-lensed samples.

As a cross-check, we compare our host magnitude to the inference of HE0435 by~\citet{Ding2017b}. They inferred the \sersic\ magnitude by fitting to the pixelized host galaxy as reconstructed by~\citet{Wong2017}. The host magnitude measured by~\citet{Ding2017b} is: $mag = 21.75 \pm 0.13$. %, $R_{eff} = 0.82 \pm 0.14$ and $n = 3.94 \pm 0.14$. 
The results are in excellent agreement with the measurement reported here, i.e., $mag = 21.50 \pm 0.35$.

@@

%In summary, considering all random and systematic effects, the uncertainty of \mstar\ relative to that measured by D20 is smaller than the estimated uncertainty in the black hole mass (0.4~dex), which dominates the overall error budget.
